\Large \bf Cyclone: Warp Speed Replication for Key Value Stores

### 2017-07-27 paper/paper.tex
More updates.

DPDK does not by itself provide a TCP stack. This is not a problem. RAFT (and
indeed most consensus protocols) tolerate network losses and reordering by
design due to the need to support asynchronous communication. In addition, most
datacenter networks rarely drop or reorder packets and provide full bisection
bandwith between servers that might serve as Cyclone replicas. We therefore
jettisoned TCP and chose to send raw IP packets encapsulated in ethernet
frames. We currently follow this communication model both for server to server
communication between replicas as well as client server communication for
requests and responses. Although a detailed evaluation is made later, switching
from the kernel TCP/IP stack to DPDK reduces the latency between machines in our
testbed from {\tt 24 us} to as low as {\tt 7 us}, providing a significant boost
to performance.

@@

Cyclone takes a software only approach to improving replication performance
using commodity network hardware. At the other end of the spectrum, Consensus in
a box~\cite{consensus_box} implements the Zookeeper atomic broadcast protocol
together with a NIC on an FPGA. That work ignored durability focusing purely on
replication performance. Latency numbers are impressive due to the FPGA that cuts
out the path from CPU caches to the NIC and are reported to be as low as 3us for
a round trip from leader to follower replica compared to the 7us we observe with
DPDK. Notably however they reported 7us when using TCP/IP from the FPGA rather
than their own connection oriented network protocol. On the other hand peak
replication throughput appears to be 6M replications/sec on 66 Gbps of aggregate
network bandwith, comparable to the mark obtained by Cyclone on an aggregate 40
Gbps of network bandwidth. We believe the general applicability afforded by our
software only packet switching solution makes it a compelling alternative to an
FPGA based solution even given the larger latency.

Network Ordered Paxos~\cite{nopaxos} reprograms top of rack switches to order
client requests and multicast them to replicas - attacking replication latency
by removing the need to direct requests through a leader replica. It is hard to
compare performance due to the 1Gbps links used to leaf nodes in NoPaxos but we
note that avoiding changes to switching infrastructure was a specific constraint
in the design of Cyclone due to the difficulty we encountered in convincing
public cloud customers to allow such changes to be made.

{CRANE}~\cite{crane} is a system for replicating mutithreaded programs using
Paxos. It uses deterministic multi-threading to ensure replicas converge to the
same state as opposed to semantic equivalence in Cyclone using the commutativity
of the key value store interface. Although CRANE is more generally applicable in
that it makes no assumptions about the application API, deterministic
multithreading comes with high overheads for applications that store large
amounts of data - CRANE reports around a 2X slowdown for MySQL likely rendering
it unsuitable as a replacement for the write ahead log in key value stores.

A number of researchers have proposed using high performance networking hardware
for improving the performance of various distributed system primitives such as
distributed transactions~\cite{farm} and replication~\cite{dare, mojim}. We note
that the ideas in Cyclone are not specific to ethernet. The same ideas of packet
switching and horizontal scaling can be equally applied to RDMA. However, we
chose commodity ethernet due to short term deployment plans that are a
priority design target.

@@

Finally, we note that scaling a log horizontally for concurrency is also an idea
adopted by filesystems. NOVA~\cite{nova} is an example that uses a per-inode log
to improve concurrency when the filesystem is placed on NVM attached to the
memory bus. Cyclone also uses a distinct NVM log per control plane core in a
similar vein but does not maintain a namespace since the number of NVM logs is a
small fixed number equal to the number of control plane cores.

>>>>

DPDK does not by itself provide a TCP stack. This is not a problem. RAFT (and
indeed most consensus protocols) tolerate network losses and reordering by
design due to the need to support asynchronous communication. In addition, most
datacenter networks rarely drop or reorder packets and provide full bisection
bandwith between servers that might serve as Cyclone replicas. We therefore
jettisoned TCP and chose to send raw IP packets encapsulated in ethernet
frames. We currently follow this communication model both for server to server
communication between replicas as well as client server communication for
requests and responses. Although a detailed evaluation is made later, switching
from the kernel TCP/IP stack to DPDK reduces the latency between machines in our
testbed from {\tt 24 us} to as low as {\tt 5 us}, providing a significant boost
to performance.

@@

Cyclone takes a software only approach to improving replication performance
using commodity network hardware. In contrast, Mojim~\cite{mojim} pairs directly
attached persistent memory with RDMA to replicate persistent data structures. On
40Gbps infiniband, Mojim reported being able to mirror data at 4GB/s. In
contrast, Cyclone can replicate data three ways using 4*10 Gbps commodity
ethernet NICs at a peak of roughly 3.7 GB/s. Notably this is while Cyclone is
also running the RAFT consensus protocol and draining the log to a high capacity
flash SSD, while Mojim is doing simple mirroring of main memory. One can
conclude that network offload via RDMA and Infiniband is not an a-priori
requirement for high throughput replication. This is made possible by Cyclone's
approach that relieves the CPU from the burden of data copies. Of course, the
latency of a hop on commodity ethernet at 5 us with our DPDK stack is larger
than the close to 2us latency with Infiniband. Depending on the use case for the
key value store application the extra 6us with ethernet for the round trip to
the replica might be a concern for some applications. We note that mapping the
commutativity of key value store interfaces to multiple logs in Cyclone is a
technique that can also be applied to RDMA. The authors of Mojim confirm that
concurrency is a significant determinant of performance when replicating over
RDMA.

Consensus in a box~\cite{consensus_box} implements the Zookeeper atomic
broadcast protocol together with a NIC on an FPGA. That work ignored durability,
focusing purely on network performance for replication. Latency numbers are
excellent due to the FPGA that cuts out the path from CPU caches to the NIC and
are reported to be as low as 3us for a round trip from leader to follower
replica compared to the 6us we observe with DPDK. Notably however they reported
7us when using TCP/IP from the FPGA rather than their own connection oriented
network protocol. On the other hand their peak replication throughput is 6M
replications/sec on 66 Gbps of aggregate network bandwith, comparable to the
same mark obtained by Cyclone on an aggregate 40 Gbps of network bandwidth. We
believe the general applicability afforded by our software only solution
together with the fact that we provide persistence by design makes it a
compelling alternative to an FPGA based solution even given the larger latency.

{CRANE}~\cite{crane} is a system for replicating mutithreaded programs using
Paxos. It uses deterministic multi-threading to ensure replicas converge to the
same state as opposed to semantic equivalence in Cyclone using the commutativity
of the key value store interface. Although CRANE is more generally applicable in
that it makes no assumptions about the application API, deterministic
multithreading comes with high overheads for applications that store large
amounts of data - CRANE reports around a 2X slowdown for MySQL, likely rendering
it unsuitable as a replacement for the write ahead log in key value stores.

@@
